---
title: "ESM 206 Lab 5"
author: "Richard Viebrock"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with "lubridate" package
- Get counts of observations with count()
- Uncount rows with the Uncount() function
- One and two-sample t-test
- Create a heat map with geom_tile()

```{r, include = FALSE}

# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
  
```

Use tidyr::uncount() function to convert our obster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm))+
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram

ggplot(lobster_tidy, aes(x = size_mm))+
  geom_histogram(aes(fill = site))+
  facet_wrap(~site,
             scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm))+
  geom_qq()+
  facet_wrap(~site)

```

#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' package to convert to date format and then to help us easily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))

```

Now let' parse year and month using lubridat::month() and lubridate::year()

```{r}

lobster_new_date <- lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
         obs_year = lubridate::year(date_new))

```

Now let's find counts of observed lobsters based on different heirarchical groupings:

First, let's count lobsters by year & month

```{r}

lobster_ym <- lobster_new_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_new_date %>% 
  count(obs_year)

lobster_site <- lobster_new_date %>% 
  count(site)

```

If we want to create a summayr table that contains statistics other than counts by group, it's easier to use group_by() + n()

```{r}

lobster_summary <- lobster_new_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

```

#### Find confidence intervals

Use t.test() function to find confidence intervals for one sample and perform t-tests to compare means of two samples (...this will be covered conceptually in lectures Week 6)

```{r}

ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)

```

#### Two-sample t-test to compare means

We want to as, is there a significant differene in lobster lengths and Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a two-sample t-test for means comparison is appropriate. 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest

```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.5).

```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)
mn_ttest2

mn_ttest2$p.value
mn_ttest$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

Here is the t-statistic `r mn_ttest2$statistic`

#### Now: a heatmap!

```{r}

lobster_ys <- lobster_new_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site))+
  geom_tile(aes(fill = n))

```

